Death by Dropdown? Engineer Insightful Shiny Apps with Behavioral Science

### 2025-09-05 conference-materials/2025_posit-conf/draft.qmd
Refactor dashboard journey and BID framework slides
*   Refactor "Your Dashboard Journey" to "Your Dashboard = A Road Trip"
*   Restructure content for better organization and visual appeal
*   Emphasize the mnemonic device for the BID framework

---
title: "Death by Dropdown? Use Behavioral Science To Engineer Insightful Shiny Apps"
author: "Jeremy Winget, PhD"
execute:
  echo: true
format:
  revealjs:
    theme: default
    css: custom.css
    slide-number: true
    transition: fade
---

@@

::: {.large}
**Ever open a Shiny app, see 18 filters across 6 tabs, and just...close your laptop?**
:::

::: {.fragment}
<div class="tenor-gif-embed" data-postid="17038619" data-share-method="host" data-aspect-ratio="1.77778" data-width="100%"><a href="https://tenor.com/view/ron-swanson-parks-and-recreation-parks-and-rec-computer-dumpster-gif-17038619">Ron Swanson Parks And Recreation GIF</a>from <a href="https://tenor.com/search/ron+swanson-gifs">Ron Swanson GIFs</a></div> <script type="text/javascript" async src="https://tenor.com/embed.js"></script>
:::

::: {.notes}
- Say it like a confession → quick laugh
- Tag: "death by dropdown → dashboard rot"
- Transition: "I saw a *human* problem, not a tech problem"
:::

@@

::: {.incremental}
- Who's built a Shiny app with 10+ filters?
- Who's had a user ask for "just one more dropdown"?
- Who's dashboard has more `selectInput()` than actual insights?
- Who's secretly proud of that 47-tab monstrosity you built in 2019?
:::

::: {.fragment}
**Welcome to Dashboard Anonymous. You're among friends.** 😅
:::

::: {.notes}
- Get audience engaged early
- Build connection through shared pain
- Light self-deprecation works well
:::

@@

:::: {.columns}
::: {.column width="60%"}
### Applied Social Psychologist → Full-Stack Developer

- **PhD in behavioral science** studying how people process information
- **Led Shiny teams** at scale (100+ dashboards)
- **Core insight:** Most dashboard failures aren't technical—they're psychological
:::

::: {.column width="40%"}
::: {.fragment}
*"As a behavioral scientist in the tech world, I've seen firsthand how a dash of psychology can rescue users from **'death by dropdown'"***
:::
:::
::::

## Your Dashboard Journey 🚗 {.center}

::: {.large}
**Building a dashboard is like planning a road trip for your users**
:::

::: {.fragment}
You wouldn't start driving without:

- Knowing your destination
- Checking the car
- Planning for weather
- Mapping the route
- Confirming arrival
:::

@@

::: {.fragment}
**Behavior Insight Design (BID)**: Not just UX advice—it's behavioral science integrated into your development workflow
:::

::: {.notes}
- We stand on UX giants' shoulders
- BID's differentiator: evidence-based psychology at each step
- Like having a behavioral scientist in the console
:::

## The Road Trip Journey 🚗

:::: {.columns}
::: {.column width="50%"}
### The 5 Stages

1. **INTERPRET** 🗺️ Plan your destination
2. **NOTICE** ⚠️ Check the dashboard warnings
3. **ANTICIPATE** 🌦️ Pack for weather changes
4. **STRUCTURE** 🛣️ Map the scenic route
5. **VALIDATE** ✅ Confirm safe arrival
:::

::: {.column width="50%"}
::: {.fragment}
### Mnemonic Device 🧠

**"Interfaces Need A Smart Vision"**

I-N-A-S-V = Your roadmap to better dashboards

*Tattoo this on your keyboard hand* 😄
:::
:::
::::

@@

## Every Shiny Developer's First Dashboard

::: {.fragment}
![](img/this-is-fine.jpg)<br>
*"But it has ALL the features!"*
:::

>>>>

---
title: "Death by Dropdown? Engineer Insightful Shiny Apps with Behavioral Science"
author: "Jeremy Winget, PhD"
execute:
  echo: true
format:
  revealjs:
    theme: default
    css: custom.css
    slide-number: true
    transition: fade
---

@@

::: {.large}
**Ever open a dashboard, see 18 filters across 6 tabs, and then...close your laptop?**
:::

::: {.fragment .large}
**Or just get the urge to...**<br>
<div class="tenor-gif-embed" data-postid="17038619" data-share-method="host" data-aspect-ratio="1.77778" data-width="100%"><a href="https://tenor.com/view/ron-swanson-parks-and-recreation-parks-and-rec-computer-dumpster-gif-17038619">Ron Swanson Parks And Recreation GIF</a>from <a href="https://tenor.com/search/ron+swanson-gifs">Ron Swanson GIFs</a></div> <script type="text/javascript" async src="https://tenor.com/embed.js"></script>
:::

::: {.notes}
:::

@@

::: {.incremental}
- Who's built a Shiny app with 10+ filters?
- Who's had a user ask for "just one more dropdown"?
- Who's dashboard has more `selectInput()` than actual insights?
:::

::: {.fragment}
**Welcome to my impromptu Dashboard Support Group. You're among friends.** 😅
:::

::: {.notes}
- Confession time...
- Who's built a one of those dashboards? ...me
- Who's been forced to build one bc the client just kept asking for one more filter? ...me
- Who's ever used selectInput more than they probably should have? ...me
- You're not alone!
:::

@@

:::: {.columns}
::: {.column width="60%"}
### Applied Social Psychologist → Full-Stack Engineer

- **PhD in Applied Psychology** studying how people process information
- **Led Shiny teams** at scale, providing enterprise dashboards
- **Core insight:** Most dashboard failures aren't technical, they're psychological
:::

::: {.column width="40%"}
::: {.fragment}
*"As a behavioral scientist in the tech world, I've seen firsthand how a dash of psychology can rescue users from that feeling of **'death by dropdown'"***
:::
:::
::::

## Your Dashboard = A Road Trip 🚗 {.center}

:::: {.columns}
::: {.column width="50%"}
### You wouldn't make a long drive without:
- **I**nterpreting travel guides
- **N**oticing any mechanical problems
- **A**nticipating roadblocks ahead
- **S**tructuring the best route
- **V**alidating everyone arrived happy
:::

::: {.column width="50%"}
::: {.fragment}
### So why build dashboards without:
- **I**nterpreting user needs
- **N**oticing friction points
- **A**nticipating behavior
- **S**tructuring a journey
- **V**alidating understanding
:::
:::
::::

::: {.notes}
- 20 seconds
- "This is the BID framework"
- Transition: "Let me show you the GPS"
:::

@@

**"Interfaces Need A Smart Vision"** - Remember: I-N-A-S-V

::: {.fragment}
**Behavioral Insight Design (BID)**: *Not just UX advice - it's behavioral science in your dev workflow*
:::

::: {.notes}
- 30-45 seconds MAX
- "Today we focus on the middle three - where psychology transforms dashboards"
- Quick transition to stages
:::

@@

## Every Dev's First Dashboard

![](img/this-is-fine.jpg)<br>
*"But it has ALL the features!"*
